---
title: "taylorswift13"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{taylorswift13}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(taylorswift13)
```



# Introduction

Welcome to **taylorswift13**! This vignette provides an overview of the package, including how to install it, access datasets, and utilize the Shiny app for data visualization and analysis.

The site is a tribute to Taylor Swift, a renowned singer-songwriter known for her captivating music and storytelling. The package contains datasets on Taylor Swift's albums and songs, enabling fans and analysts to explore her discography in-depth.




# Installation

To install **taylorswift13** from GitHub, use the following commands:

```R
install.packages("remotes")
remotes::install_github("ETC5523-2024/taylorswift13")
```

# Accessing Datasets

taylorswift13 includes three main datasets:

- taylor_album: Information about Taylor Swift's albums.
- taylor_album_songs: Details of songs from each album.
- taylor_all_songs: Comprehensive data on all Taylor Swift's songs.

To access these datasets, use the following commands:

```R
data("taylor_album")
data("taylor_album_songs")
data("taylor_all_songs")
```

# Loading and Exploring Datasets

taylor_album
```R
data(taylor_album)
head(taylor_album)
```
Description: This dataset provides metadata about Taylor Swift's albums, including release dates and ratings.

# taylor_album_songs
```R
data(taylor_album_songs)
head(taylor_album_songs)
```

Description: Detailed information about each song within Taylor Swift's albums, covering musical features and release details.


# taylor_all_songs
```R
data(taylor_all_songs)
head(taylor_all_songs)
```

Description: An aggregated dataset encompassing all songs by Taylor Swift, facilitating comprehensive analyses across her entire discography.

# Shiny App

```R
launch_app()
```

The Shiny app allows you to visualize various aspects of Taylor Swift's discography, including song characteristics, album metrics,and song comparisons. Explore the app to gain insights into Taylor Swift's musical evolution!

Features of the Shiny App:

- Dashboard Layout: Organized interface for easy navigation.
- Interactive Plots: Utilize ggplot2 and plotly for dynamic visualizations.
- Data Tables: Explore datasets with sortable and searchable tables using DT.
- Filter Options: Customize views based on album, song features, and more.

# Example

Here's a simple example of how to visualize the distribution of song energy across Taylor Swift's albums using ggplot2:

```R
library(taylorswift13)
library(ggplot2)

# Load the album songs data
data(taylor_album_songs)

# Plot energy distribution
ggplot(taylor_album_songs, aes(x = album_name, y = energy)) +
  geom_boxplot(fill = "steelblue") +
  theme_minimal() +
  labs(title = "Distribution of Song Energy Across Albums",
       x = "Album",
       y = "Energy Score")

```

# Conclusion

We hope `taylorswift13` enhances your analysis and appreciation of Taylor Swift's musical journey! For more information, check out the package documentation and explore the Shiny app.

